---
title: "Mediating traits in Attie DO data"
author: "Frederick J. Boehm"
date: "2/12/2020"
output: html_document
params:
  lastmod: !r paste0(lubridate::now(), " (", Sys.timezone(), ")")
---

Last modified: `r params$lastmod`.

## Overview

We'd like to clarify mediation methods and explore them in the context of the Attie DO data. Basically, I want to examine methods for multivariate mediation analysis with the "difference" method and compare them with the methods of @chick2016defining.

It seems that the methods that vanderWeele and others use differ in important ways from the LOD difference approach that @chick2016defining use.

I want to ensure that I understand these differences and think about which I should use for future analyses.

We'll use linear regression methods with the only variance component being that of the random error term -- so, no "polygenic random effect" related to the kinship matrix.



## Download Attie data

```{r}
fn <- "../data/raw_data/attie.RData"
if (!file.exists(fn)) download.file(url = "https://datadryad.org/stash/downloads/file_stream/63306", destfile = fn)
```

```{r}
genofn <- "../data/derived_data/attie_geno.rds"
kinsfn <- "../data/derived_data/attie_kinship.rds"
mapfn <- "../data/derived_data/attie_map.rds"
exprfn <- "../data/derived_data/attie_expr.rds"
covafn <- "../data/derived_data/attie_covar.rds"
if (!file.exists(genofn)){
  load(fn)
  saveRDS(genoprobs, genofn)
  saveRDS(K, kinsfn)
  saveRDS(map, mapfn)
  saveRDS(dataset.islet.rnaseq$expr, exprfn)
  saveRDS(dataset.islet.rnaseq$covar, covafn)
}
geno <- readRDS(genofn)
kinship <- readRDS(kinsfn)
map <- readRDS(mapfn)
expr <- readRDS(exprfn)
covar <- readRDS(covafn)

```


## Read the peaks annotations and fit linear models at the peaks.

What do we want to record for each model fit? 

The coefficient esimates, and RSS?? It seems that we should save all of the output from calls to `lm`.

First, choose a superstrong peak.

For every trait-mediator-peak triple, we need to fit 4 models: with and without genotypes, with and without putative mediator, for a single trait peaks. However, we should also fit the models that have the outcome variable being the mediator. So, actually, we need more than four models per triple.

## Questions

1. What are the meanings of direct and indirect effects? Are they meaningful in systems genetics studies?

1. Which of the causal inference (absence of confounding) assumptions holds in the systems genetics setting? I believe that Vanderweele's text outlines four such assumptions. Check each of them.



1. 














## References

